---
title: "Text Analysis of Pop Lyrics: Sentiment, Word Frequency"
author: "Carson Zhang (cxzhang4)"
date: "7/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tidytext)
```

```{r}
artist_albums = read_csv("artist_albums.csv")
pop_lyrics = read_csv("pop_lyrics.csv")
pop_lyrics_by_word = pop_lyrics %>%
  unnest_tokens(word, lyric)
```

```{r}
lyrics_afinn = pop_lyrics_by_word %>%
  inner_join(get_sentiments("afinn")) %>%
  group_by(index = track_title, album, artist, gender) %>%
  summarise(sentiment = sum(value)) %>%
  mutate(method = "AFINN")

lyrics_bing_and_nrc = bind_rows(pop_lyrics_by_word %>%
                               inner_join(get_sentiments("bing")) %>%
                               mutate(method = "Bing"),
                             pop_lyrics_by_word %>% 
                               inner_join(get_sentiments("nrc")) %>%
                               filter(sentiment %in% c("positive", "negative")) %>%
                               mutate(method = "NRC")) %>%
  count(method, index = track_title, album, artist, gender, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>%
  mutate(sentiment = positive - negative)
```

```{r}
lyrics_sentiments = bind_rows(lyrics_afinn, lyrics_bing_and_nrc)
```

```{r}
lyrics_afinn_p = lyrics_sentiments %>% 
  filter(method == "AFINN") %>% 
  pivot_wider(names_from = method, values_from = sentiment)

lyrics_bing_p = lyrics_sentiments %>% 
  filter(method == "Bing") %>% 
  pivot_wider(names_from = method, values_from = sentiment)

lyrics_nrc_p = lyrics_sentiments %>% 
  filter(method == "NRC") %>% 
  pivot_wider(names_from = method, values_from = sentiment)

lyrics_sentiments_p = lyrics_afinn_p %>%
  inner_join(lyrics_bing_p, by = c("index", "artist", "album")) %>%
  inner_join(lyrics_nrc_p, by = c("index", "artist", "album"))

lyrics_sents_p_c = lyrics_sentiments_p %>%
  select(index, artist, album, gender, AFINN, Bing, NRC) %>%
  ungroup()

# write_csv(lyrics_sents_p_c, "lyrics_sents_p_c.csv")
```

```{r}
artist_song_counts = lyrics_sents_p_c %>%
  count(artist)
```

```{r}
lyrics_sents_totals = lyrics_sents_p_c %>%
  group_by(artist) %>%
  summarise(AFINN_total = sum(AFINN), Bing_total = sum(Bing), NRC_total = sum(NRC))
```

```{r}
artist_song_avgs = lyrics_sents_totals %>%
  inner_join(artist_song_counts, by = "artist") %>%
  mutate(AFINN_song_avg = AFINN_total / n,
         Bing_song_avg = Bing_total / n,
         NRC_song_avg = NRC_total / n,
         n_songs = n)

scale_col = function(x, na.rm = FALSE) {
  (x - mean(x, na.rm = na.rm)) / sd(x, na.rm)
}

artist_song_avgs_scaled = artist_song_avgs %>% 
  mutate(across(ends_with("song_avg"), scale_col))
```

```{r}
artist_song_avgs_scaled %>%
  ggplot(aes(AFINN_song_avg)) +
  geom_histogram()
```

```{r}
# artist_song_avgs_scaled %>%
#   mutate(AFINN_cdf = cume_dist(AFINN), 
#          Bing_cdf = cume_dist(Bing), 
#          NRC_cdf = cume_dist(NRC)) %>%
#   filter(AFINN_cdf < 0.05 & Bing_cdf < 0.05 & NRC_cdf < 0.05)
```

```{r}
lyrics_sents_p_c_scaled = lyrics_sents_p_c %>%
  mutate(across(c("AFINN", "Bing", "NRC"), scale_col)) %>%
  mutate(AFINN_cdf = cume_dist(AFINN), 
         Bing_cdf = cume_dist(Bing), 
         NRC_cdf = cume_dist(NRC))
```

So how do we define a very negative song?

2 standard deviations below the mean is a reasonable cutoff value for "very negative": in any case, it is noticeably more negative than the average song.

We include all songs who meet this cutoff on any one of the sentiment measures: this means that there is some interpretation of the song as very negative, even if other measures don't agree.

```{r}
very_neg_songs_by_sd = lyrics_sents_p_c_scaled %>%
  filter(AFINN < -2 | Bing < -2 | NRC < -2)

(very_neg_songs_by_sd %>% nrow()) %>%
  "/"(lyrics_sents_p_c_scaled %>% nrow())
```

```{r}
very_neg_songs_by_cdf = lyrics_sents_p_c_scaled %>%
  filter(AFINN_cdf < 0.1 | Bing_cdf < 0.1 | NRC_cdf < 0.1)

(very_neg_songs_by_cdf %>% nrow()) %>%
  "/"(lyrics_sents_p_c_scaled %>% nrow())
```

R tip: infix operators are functions! That's why we can use the "/" operator in conjunction with the pipe.

```{r}
artist_counts_cdf = very_neg_songs_by_cdf %>%
  count(artist)
artist_counts_sd = very_neg_songs_by_sd %>%
  count(artist)

artist_props = artist_counts_cdf %>%
  inner_join(artist_song_counts, by = "artist") %>%
  rename(very_neg_cdf = n.x, total_songs = n.y) %>%
  inner_join(artist_counts_sd, by = "artist") %>%
  rename(very_neg_sd = n) %>%
  mutate(cdf_prop = very_neg_cdf / total_songs, sd_prop = very_neg_sd / total_songs)
```

```{r}
highlight_artist = function(df, artist_name) {
  df %>% 
    mutate(highlight = ifelse(artist == artist_name, "yes", "no")) %>%
    return()
}

artist_highlight_bar = function() {
  list(geom_col(),
       coord_flip(),
       scale_fill_manual(values = c("yes" = "tomato", "no" = "gray"),
                         guide = FALSE)
  )
}
```

```{r}
artist_props %>%
  highlight_artist("Taylor Swift") %>%
  ggplot(aes(x = reorder(artist, cdf_prop), y = cdf_prop, fill = highlight)) +
  artist_highlight_bar() +
  ggtitle("title")
```

```{r}
artist_props %>%
  highlight_artist("Taylor Swift") %>%
  ggplot(aes(x = reorder(artist, sd_prop), y = sd_prop, fill = highlight)) +
  artist_highlight_bar() +
  ggtitle("title")
```

## Which are the most [insert artist here] words?

```{r}
# library(janeaustenr)
# 
# book_words <- austen_books() %>%
#   unnest_tokens(word, text) %>%
#   count(book, word, sort = TRUE)

# book_words
```

```{r}
artist_word_counts = pop_lyrics_by_word %>%
  count(artist, word, sort = TRUE)

total_words = artist_word_counts %>%
  group_by(artist) %>%
  summarise(total = sum(n))

artist_word_counts = left_join(artist_word_counts, total_words)

artist_word_counts
```

```{r}
artist_word_counts %>%
  bind_tf_idf(word, artist, n)
```

```{r}
artist_word_counts %>%
  arrange(tf_idf)
```

```{r}
album_word_counts = pop_lyrics_by_word %>%
  count(album, artist, word, sort = TRUE)

total_album_words = album_word_counts %>%
  group_by(artist) %>%
  summarise(total = sum(n))

album_word_counts = left_join(album_word_counts, total_words) %>%
  bind_tf_idf(word, album, n)
```

```{r}
album_word_counts %>% 
  filter(artist == "Taylor Swift") %>%
  arrange(desc(tf_idf)) %>%
  mutate(word = factor(word, levels = rev(unique(word)))) %>% 
  group_by(album) %>% 
  top_n(5) %>% 
  ungroup() %>%
  ggplot(aes(word, tf_idf, fill = album)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  facet_wrap(~album, ncol = 2, scales = "free") +
  coord_flip()
```

```{r}
album_word_counts %>%
  filter(album == "Folklore") %>%
  # arrange(desc(tf_idf)) %>%
  slice_max(tf_idf, n = 20) %>%
  ggplot(aes(reorder(word, tf_idf), tf_idf, fill = album)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  coord_flip()
```

```{r}
ts_word_counts = pop_lyrics_by_word %>%
  filter(artist == "Taylor Swift") %>%
  count(album, artist, word, sort = TRUE)

total_ts_words = album_word_counts %>%
  group_by(artist) %>%
  summarise(total = sum(n))

ts_word_counts = left_join(ts_word_counts, total_ts_words) %>%
  bind_tf_idf(word, album, n)
```

```{r}
ts_word_counts %>%
  filter(album == "Folklore", word %in% c("bitch", "shit", "fuck")) %>%
  # arrange(desc(tf_idf)) %>%
  slice_max(tf_idf, n = 40) %>%
  ggplot(aes(reorder(word, tf_idf), tf_idf, fill = album)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  coord_flip()
```

Further work:

- Shiny app to display all of this for any artist/album in dataset
- Entity-level sentiment analysis to determine how negatively an artist refers to other people
- Figure out which songs concern romantic relationships
- Better statistical tests for gender differences
